Connecting the Popularity Adjusted Block Model to the Generalized Random Dot Product Graph for Community Detection and Parameter Estimation

### 2021-07-15 summary.tex
minor edits

\begin{itemize}
\tightlist
\item
  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
\item
  Number of underlying communities \(K = 2, 3, 4\)
\item
  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\), (i.e.,
  each community label has an equal probability of being drawn)
\item
  Community labels
  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
\item
  Within-group popularities
  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
\item
  Between-group popularities
  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
  \(k \neq l\)
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
\item
  Number of underlying communities \(K = 2, 3, 4\)
\item
  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
  for \(k = 1, ..., K\)
\item
  Community labels
  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
\item
  Within-group popularities
  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
\item
  Between-group popularities
  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
  \(k \neq l\)
\end{itemize}

@@

We again see community detection error trending to 0 for OSC, as well as
for SSC when \(K > 2\) (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3
continues to see \(n^{-1/2}\) decay in parameter estimation error; 
however, the MLE-based estimators for the popularity vectors decay at a slower 
rate (Fig. \ref{fig:lambda_est_k_imba}). 
The reduced rate of decay is removed 
if we reconstruct the edge probability blocks 
by taking the outer products of the popularity vectors.

@@

Comparing against the ground truth species labels, OSC achieves an
accuracy of 63\% and an adjusted Rand index of 73\%. In comparison,
\citet{noroozi2019estimation} achieved an adjusted Rand index of
73\% using sparse subspace clustering on the same dataset.

@@

In the third example, we consider the Karantaka villages data studied by
\citet{DVN/U3BIHX_2013}. For this example, we chose the \texttt{visitgo}
networks from villages 12, 31, and 46 at the household level. The label
of interest is the religious affiliation. The networks were truncated to
religions ``1'' and ``2'', and vertices of degree 0 were removed. The results are displayed in Fig. \ref{fig:households-figure} and Table \ref{tab:households-table}.

>>>>

\begin{itemize}
\tightlist
\item
  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
\item
  Number of underlying communities \(K = 2, 3, 4\)
\item
  Mixture parameters \(\alpha_k = 1 / K\) for \(k = 1, ..., K\), (i.e.,
  each community label has an equal probability of being drawn)
\item
  Community labels
  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
\item
  Within-group popularities
  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
\item
  Between-group popularities
  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
  \(k \neq \ell\)
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Number of vertices \(n = 128, 256, 512, 1024, 2048, 4096\)
\item
  Number of underlying communities \(K = 2, 3, 4\)
\item
  Mixture parameters \(\alpha_k = \frac{k^{-1}}{\sum_{\ell=1}^K \ell^{-1}}\)
  for \(k = 1, ..., K\)
\item
  Community labels
  \(z_k \stackrel{\text{iid}}{\sim} \Multinomial(\alpha_1, ..., \alpha_K)\)
\item
  Within-group popularities
  \(\lambda^{(kk)} \stackrel{\text{iid}}{\sim} \Betadist(2, 1)\)
\item
  Between-group popularities
  \(\lambda^{(k \ell)} \stackrel{\text{iid}}{\sim} \Betadist(1, 2)\) for
  \(k \neq \ell\)
\end{itemize}

@@

We again see community detection error trending to 0 for OSC, as well as
for SSC when \(K > 2\) (Fig. \ref{fig:clust_err_ct_sim_imba}). Alg. 3
continues to see \(n^{-1/2}\) decay in parameter estimation error; 
however, the MLE-based estimators for the popularity vectors decay at a slower 
rate (Fig. \ref{fig:lambda_est_k_imba}). 
The reduced rate of decay is removed 
if we reconstruct the edge probability blocks 
by taking the outer products of the popularity vectors 
(Fig. \ref{fig:lambda_est_p_imba}).

@@

Comparing against the ground truth species labels, OSC achieves an adjusted Rand index of 92\%, while SSC on the ASE achieves an adjusted Rand index of 96\%. In comparison, SSC on the adjacency matrix yields an adjusted Rand index of 73\% \citep{noroozi2019estimation}.

@@

In the third example, we analyzed the Karantaka villages data studied by
\citet{DVN/U3BIHX_2013}. We chose the \texttt{visitgo}
networks from villages 12, 31, and 46 at the household level. The label
of interest is the religious affiliation. The networks were truncated to
religions ``1'' and ``2'', and vertices of degree 0 were removed. 
The results are displayed in Fig. \ref{fig:households-figure} and 
Table \ref{tab:households-table}.
